---
title: "Reproducible Research - Peer Assessment 1"
output: html_document
---

Wen Hao Wong

### A. Loading and preprocessing data

1. Loading the data

```{r}
data <- read.csv("activity.csv", header = TRUE)
summary(data)
str(data)
head(data)
tail(data)
```

2. Through looking at the summary and structure of the data, as well as examining some rows of the dataset, it appears that no further processing/transformation is necessary at this point in time.

### B. What is mean total number of steps taken per day?

1. Making a histogram of the total number of steps taken each day

```{r}
steps_day <- aggregate(steps ~ date, data = data, FUN = sum)
barplot(steps_day$steps, ylab = "Total number of steps taken", xlab = "Date", names.arg = steps_day$date)
```

2. The **mean** and **median** total number of steps taken each day are as follows:

```{r}
mean(steps_day$steps)
median(steps_day$steps)
```

### C. What is the average daily activity pattern?

1. Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_interval <- aggregate(steps ~ interval, data = data, FUN = mean)
plot(steps_interval, type = "l", ylab = "Average number of steps taken", xlab = "Interval")
```

2. The 5-min interval that contains the maximum number of steps is as follows:

```{r}
steps_interval$interval[which.max(steps_interval$steps)]
```

### D. Imputing missing values

1. The total number of missing values in the dataset is as follows:

```{r}
sum(is.na(data))
```

2. My strategy is to use the means of the 5-minute intervals as fillers for the missing values.

3. Using this strategy, I create a new dataset but with the mising data filled:

```{r}
data_merged <- merge(data, steps_interval, by = "interval", suffixes = c("",".y"))
nas <- is.na(data_merged$steps)
data_merged$steps[nas] <- data_merged$steps.y[nas]
data_filled <- data_merged[, c(1:3)]
```

4. I then make a histogram of the total number of steps taken each day. The **mean** is the same but the **median** is (a little) different. The impact of imptung mising values seems to be small.

```{r}
steps_day_filled <- aggregate(steps ~ date, data = data_filled, FUN = sum)
barplot(steps_day_filled$steps, ylab = "Total number of steps taken", xlab = "Date", names.arg = steps_day_filled$date)
mean(steps_day_filled$steps)
median(steps_day_filled$steps)
```

### E. Are there differences in activity patterns between weekdays and weekends?

1. Creating a new factor variable in the dataset with two levels ??? ???weekday??? and ???weekend??? indicating whether a given date is a weekday or weekend day:

```{r}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
data_filled$daytype <- as.factor(sapply(data_filled$date, daytype))
```

2. Making a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):

```{r}
par(mfrow = c(2,1))
for (type in c("weekend", "weekday")) {
    steps_interval_filled_type <- aggregate(steps ~ interval, data = data_filled, subset = data_filled$daytype==type, FUN = mean)
    plot(steps_interval_filled_type, type = "l", main = type, ylab = "Average number of steps taken", xlab = "Interval")
}
```